---
title: "Final Project"
author: "Ronald Tinashe Nhondova - RTN6"
date: "11/12/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(CASdatasets)
library(ggplot2)
library(mice)
library(dplyr)
library(VIM)
library(lattice)
library(tweedie)
```

```{r data_import}
data(brvehins1a,brvehins1b,brvehins1c,brvehins1d,brvehins1e)
```

```{r Initial EDA}
summary(brvehins1a)
summary(brvehins1b)
summary(brvehins1c)
summary(brvehins1d)
summary(brvehins1e)
```

```{r visualize missing data}
aggr(brvehins1a,col=c("lightblue3","darkred"),numbers=TRUE,sortVars=TRUE,labels=names(brvehins1a),
                    cex.axis=.7,gap=3,ylab=c("Proportion missing","Missingness pattern"))

aggr(brvehins1a[which(brvehins1a$Gender == "Corporate"),],col=c("lightblue3","darkred"),numbers=TRUE,sortVars=TRUE,labels=names(brvehins1a),
                    cex.axis=.7,gap=3,ylab=c("Proportion missing","Missingness pattern"))
```

# Data adjustments
* Policies with Gender ***Corporate*** and missing m

```{r data wrangling}
brvehins_comb <- rbind_all(list(brvehins1a,brvehins1b,brvehins1c,brvehins1d,brvehins1e))

brvehinsClaims <- brvehins_comb %>% mutate(ClaimsTotal=rowSums(brvehins_comb[,c(19:23)])) %>% select(-c(StateAb,ExposFireRob,PremFireRob,VehModel,ClaimNbRob,ClaimNbPartColl,ClaimNbTotColl,ClaimNbFire,ClaimNbOther,ClaimAmountRob,ClaimAmountPartColl, ClaimAmountTotColl, ClaimAmountFire, ClaimAmountOther))

aggr(brvehinsClaims,col=c("lightblue3","darkred"),numbers=TRUE,sortVars=TRUE,labels=names(brvehinsClaims),
                    cex.axis=.7,gap=3,ylab=c("Proportion missing","Missingness pattern"))

```

```{r impute-missing-data}
brvehinsClaims_imp <- mice(brvehinsClaims,m=2,defaultMethod=c("norm","logreg","polyreg","polr"),print=F,seed = 1234)
stripplot(brvehinsClaims_imp, col=c("grey","darkred"),pch=c(1,20))
densityplot(brvehinsClaims_imp)
```

```{r data-wrangling-2}

brvehinsClaims_om <- brvehinsClaims %>% filter((Gender == "Corporate" & rowSums(is.na(brvehinsClaims[,c(4:10)]))==0) | (Gender != "Corporate" & rowSums(is.na(brvehinsClaims))==0))

brvehinsClaimSev <- brvehins_comb %>% select(-c(ExposFireRob,PremFireRob,VehModel)) %>% filter((Gender == "Corporate" & rowSums(is.na(brvehins_comb[,c(2:20)]))==0) | (Gender != "Corporate" & rowSums(is.na(brvehins_comb))==0)) %>% mutate(claims_SevRob=ClaimAmountRob/ClaimNbRob,claims_SevPartColl=ClaimAmountPartColl/ClaimNbPartColl,claims_SevTotColl=ClaimAmountTotColl/ClaimNbTotColl,claims_SevFire=ClaimAmountFire/ClaimNbFire,claims_SevOther=ClaimAmountOther/ClaimNbOther) %>% filter(claims_SevRob != Inf & claims_SevPartColl != Inf & claims_SevTotColl != Inf & claims_SevFire != Inf & claims_SevOther != Inf)

summary(brvehinsClaimFreq)
summary(brvehinsClaimSev)
```

```{r remove-missing}
brvehinsClaims_nm <- brvehinsClaims %>% filter(rowSums(is.na(brvehinsClaims))==0)
summary(brvehinsClaims_nm)

brvehinsClaims_nm_imp <- mice(brvehinsClaims_nm,m=2,defaultMethod=c("norm","logreg","polyreg","polr"),print=F,seed = 1234,nnet.MaxNWts = 2000)

```

```{r modelling da}

```{r EDA part1}

ggplot(brvehinsClaims_nm,aes(x=PremTotal)) + geom_density() + scale_x_log10() + labs(title = "Premium Total")
ggplot(brvehinsClaims_nm[which(brvehinsClaims$ClaimsTotal>100000),],aes(x=ClaimsTotal)) + geom_histogram() + labs(title = "Claims Total") 



#ggplot(brvehinsClaimSev,aes(x=claims_SevRob)) + geom_histogram(bins = 40) + labs(title = "Claim Severity Robbery")
#ggplot(brvehinsClaimSev,aes(x=claims_SevTotColl)) + geom_histogram(bins = 20) + labs(title = "Claim Severity Total Collision")
#ggplot(brvehins,aes(x=ClaimNbPartColl)) + geom_bar() + labs(title = "Claim number Partial Collision")
#ggplot(brvehins,aes(x=ClaimNbTotColl)) + geom_bar() + labs(title = "Claim number Total Collision")
#ggplot(brvehins,aes(x=ClaimNbFire)) + geom_bar() + labs(title = "Claim number Fire")
#ggplot(brvehins,aes(x=ClaimNbOther)) + geom_bar() + labs(title = "Claim number Other")


```

```{r frequency-tables}
breaks = seq(0, 5000, by=500)
ClaimTotal.cut = cut(brvehinsClaims_nm$ClaimsTotal, c(-Inf,0,0.5,500,1000,1500,2000,2500,3000,10000,100000,1000000), right=FALSE)
ClaimTotal.freq = table(ClaimTotal.cut)
ClaimTotal.freq
```


```{r model initial}

```
